---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group? You will need to use `case_when` to make this work and there's an example of this in the book chapter on mutate.
**A1** [write your answer here]
1 under_10k   9310
2 10k_50k   15372
3 50k_100k    3978
4 100k_500k   4152
5 500k_1m   540
6 1m_plus   356
```{r}
# Write the code to produce your answer here
wv_with_loan_quantized <- west_virginia_ppp %>%
  mutate(
    loan_group = case_when(
      amount < 10000 ~ "under_10k",
      amount >= 10000   & amount < 50000   ~ "10k_50k", 
      amount >= 50000  & amount < 100000  ~ "50k_100k", 
      amount >= 100000  & amount < 500000  ~ "100k_500k", 
      amount >= 500000  & amount < 1000000 ~ "500k_1m", 
      amount >= 1000000  ~ "1m_plus"
      )
  ) %>% select(id,name,amount,loan_group,everything())

wv_loangroup_summary <-
  wv_with_loan_quantized %>%
  group_by(loan_group) %>%
  summarise(
    count_loans = n(),
  ) %>%
arrange(loan_group)
```

**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
Select those records with undisbursed_amount > 0
Group by servicing_lender
count the number of loans
Sort descending by loan count

**A2** [write your answer here]

Harvest Small Business Finance, LLC	139	2662012.00		
Capital Plus Financial, LLC	80	1169688.00		
Benworth Capital	70	1174336.00		
Customers Bank	52	914208.00		
Leader Bank, National Association	52	559893.00		
BSD Capital, LLC dba Lendistry	49	894588.00		
Prestamos CDFI, LLC	33	372407.00		
Davis Trust Company	32	360613.65		
Natural Capital Investment Fund, Inc.	30	1210289.66	
First United Bank & Trust	15	1701670.66	
```{r}
# Write the code to produce your answer here
undisbursed_wv <- west_virginia_ppp %>% filter(undisbursed_amount > 0)
undisbursed_wv %>%
  group_by(servicing_lender_name) %>%
  summarise(
    count_loans = n(),
    total_undisbursed = sum(undisbursed_amount)
  ) %>%
  arrange(desc(count_loans))
  

```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** [write your answer here]
These are all questionable "fintech" business partnerships that may have attempted to leverage the PPP program in ways that are outside the law. This NYT article https://www.nytimes.com/2021/06/27/business/ppp-relief-loans-blueacorn-womply.html discusses how online businesses partnered with some of these lenders to mass-process loan applications. However, it also appears that there have been numerous fraudulent applications (in that the companies/lenders applied on behalf of businesses that were unaware that they were applying for PPP loans) or that loans were approved but not disbursed. Since lenders were compensated based on the number and volume of loans, these businesses benefitted form the processing fees charged, but were not serving their applicants. see https://www.usatoday.com/story/news/investigations/2021/08/17/covid-19-ppp-loan-borrower-fraud-fueled-fintechs-report-finds/8124023002/ as well for additional details.
**Q4** Which month in 2021 had the most loans approved?
filter out 2021 loans
group by month
count the number of loans approved
**A4** [write your answer here]
2021-03-01	4404	
March 2021 with 4404 loans
```{r}
# Write the code to produce your answer here
wv_2021 <- west_virginia_ppp %>% filter(floor_date(date_approved, "year") == date("2021-01-01"))
wv_2021 %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
  
```
