---
title: "lab_12"
author: "derek willis"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key
* Our usual libraries, plus the geographic ones

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)
library(stringi)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggrepel)
library(sf)
library(tidycensus)
library(tigris)
options(tigris_use_cache = TRUE)

# Instantiate the census key
mykey <- scan("../../censusapikey.txt", what=character())
census_api_key(mykey, install=TRUE, overwrite=TRUE)
```

## Questions

**Q1.** You are interested in tracking vaccinations by age group, and in particular how many children are being vaccinated over time. You want to generate a sentence that describes the daily change in first daily dose among children between 5-11 years old, and also compares it to the most recent same day of the week, describing the change in percentage terms and the direction of the change.

Using [CSV data posted by the state](https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/MD_COVID19_VaccinationPercentAgeGroupPopulation.csv), making the column names more readable and ensuring each day is formatted as a date column that ONLY contains the date, not a time.

Then write code that will construct the sentence about the data described above comparing the most recent data to the previous date and the most recent previous date for the same day of the week. To do that, you will need:

1. A dataframe with vaccination data for the under-12 age group.
2. A column in your dataframe that contains the day of the week that each date represents (you may need to google how to do that).
3. Data from the previous day and the most recent previous same day of the week.
4. The percentage changes compared to the previous day and most recent previous same day of the week.
5. A template for your sentence using paste.

Put your sentence in the answer below.

**A1.**
"The number of children under age 12 in the State of Maryland having their first vaccine dose on Wednesday, November 17 decreased by -49.99 percent compared to the day before, and decreased by -70.75 percent compared to the week before."

```{r}
vaccinations_by_age_group <- read_csv("https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/MD_COVID19_VaccinationPercentAgeGroupPopulation.csv")

# tinkering to see how this works...
#format(mdy_hms("10/30/2021 15:00:00"), "%A")

vaccinations_under_12 <- vaccinations_by_age_group %>%
  filter(str_detect(AgeRange, "11 and Under")) %>%
  # Get the day of the week using format
  mutate(DayOfWeek = format(mdy_hms(VACCINATION_DATE), "%A")) %>%
  # add a column with a formatted date without the hours and minutes
   mutate(vaccination_date = date(mdy_hms(VACCINATION_DATE))) %>%
  # sort in reverse date order
  arrange(desc(vaccination_date))

#Pull out the latest record
today_vax <- vaccinations_under_12 %>% slice(1)

# This is for the day prior to the latest
yesterday_vax <- vaccinations_under_12 %>% filter(vaccination_date == today_vax$vaccination_date - 1)

# This is for one week prior to the latest
one_week_ago <-  vaccinations_under_12 %>% filter(vaccination_date == today_vax$vaccination_date - 7)

# Add the percentage change from yesterday's first dail does and the percentage change from last week's first daily dose
today_vax <- today_vax %>%
  mutate(pct_change_day = (FirstDailyDose - yesterday_vax$FirstDailyDose)/yesterday_vax$FirstDailyDose *100) %>%
  mutate(pct_change_week = (FirstDailyDose - one_week_ago$FirstDailyDose)/one_week_ago$FirstDailyDose *100)

changetext1 <- round(today_vax$pct_change_day[[1]], digits=2)
direction1 <- if_else(changetext1 > 0, "increased", "decreased")

changetext2 <- round(today_vax$pct_change_week[[1]], digits=2)
direction2 <- if_else(changetext2 > 0, "increased", "decreased")


sub <- paste("The number of children under age 12 in the State of Maryland having their first vaccine dose on ", today_vax$DayOfWeek, ", ", format(today_vax$vaccination_date, format="%B %d"), " ", direction1, " by ", changetext1, " percent compared to the day before, and ", direction2, " by ", changetext2, " percent compared to the week before.", sep="")
sub
```

**Q2.** Maryland produces a [CSV file](https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/Vax_ZIP_CODE_FirstSecondSingle_Totals_Raw.csv) showing the total number of vaccinations by zip code and -- interestingly -- it includes non-Maryland zip codes (as well as some zip codes that seem to be bad data). write a script that reads the CSV file, making the columns more readable and adding today's date. Then output the top 25 Maryland zip codes and top 25 non-Maryland zip codes based on number of people fully vaccinated, and write both dataframes out to CSV files, using the following name formats:

Maryland zips: maryland_zips_yyyymmdd.csv
Non-Maryland zips: non_maryland_zips_yyyymmdd.csv

Where yymmdd is from today's date. You MUST do this programmatically so that it can run every day without being updated. Be sure to add those two files to your GitHub repository.

To do that, you will need to:

1. Get rid of rows that do not have data for fully vaccinated people
2. Remove any rows where the zip code is less than five digits (there are several ways you could do this).
3. Figure out the "lowest" and "highest" Maryland zip codes.

Google the top 2-3 zip codes from both Maryland and non-Maryland results and describe them in your answer.

**A2.**
In Maryland: 20906, 20878, 20874. Highly populatted areas?
Out of maryland: 20011, 20009, 20002. Is it possible these are MD residents vaccinated in the listed Zip codes?

```{r}
today_date <- today(tzone = "")

# Read in all the rercords, get rid of ones with malformed zips (probably ones with leading 0!!!)
all_by_zip <- read_csv("https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/Vax_ZIP_CODE_FirstSecondSingle_Totals_Raw.csv") %>%
  mutate(vaccination_date = today_date) %>%
  mutate(ziplen = str_length(ZIP_CODE)) %>% 
  filter(ziplen == 5) %>%
  mutate(ZIP_CODE = as.character(ZIP_CODE))

# Get the population data for MD zip codes (for later, but has a list of MD zip codes!)
acs1 <- load_variables(2019, "acs5", cache = TRUE)

md <- get_acs(geography = "zcta",
              variables = c(population="B01001_001"),
              state = "MD") %>%
  mutate(ZIP_CODE = GEOID)

#
# join by ZIP_CODE
# Note that we only have GEOID values for the Maryland Zip Codes, so we can get rid of the ones where that
# value is NA
#
md_by_zip2 <- all_by_zip %>% left_join(md, by=c("ZIP_CODE"="ZIP_CODE")) %>%
  filter(!is.na(GEOID)) %>%
  arrange(desc(FullyVaccinated))

# got a list of MD zip codes from https://www.zipdatamaps.com/list-of-zip-codes-in-maryland.php
# This was not really right so I commented it out
# md_lowest_zip <- "20601"
# md_highest_zip <- "21930"
# 
# md_by_zip <- all_by_zip %>%
#   filter(stri_cmp_le(ZIP_CODE, md_highest_zip)) %>%
#   filter(stri_cmp_ge(ZIP_CODE, md_lowest_zip)) %>%
#   arrange(desc(FullyVaccinated))

others_by_zip <- all_by_zip %>% left_join(md, by=c("ZIP_CODE"="ZIP_CODE")) %>%
  filter(is.na(GEOID)) %>%
  arrange(desc(FullyVaccinated))

#stri_cmp_gt("20902", md_lowest_zip)
md_by_zip2 %>% head(25)
others_by_zip %>% head(25)

```

**Q3.** Maryland has had one of the more successful vaccination rates, but you want to understand where a significant portion of the population remains less than fully vaccinated. Make a dataframe of all Maryland zip codes from the vaccination data you used in Q2 and join it to Maryland zip code population data from the Census API. Calculate the percentage of fully vaccinated for each zip code and then make a map showing the zip codes with less than 50 percent fully vaccinated. Put a blank map with the outline of Maryland counties (which you can get from Tigris) behind it for context. Add a label to the map describing what the data is and include the date from the vaccination data (programmatically, so you can run it every day).

You WILL need to ensure that the columns you use to join the two dataframes are the same datatype and remember that join order matters.

Describe the output of the map below, and say what you think the most interesting areas are.

**A3.**
Looks like a lot of the western MD Zip codes and a bunch along the Delaware border. These might be relatively low-population areas.
```{r}
# Write code here to load the ACS variables
# I did this for the answer to #2 see above

# Calculate the percent fully vaccinated and filter out those that have less than 50%
md_by_zip_lt50 <- md_by_zip2 %>%
  mutate(pctFullVaccinated = (FullyVaccinated/estimate) * 100) %>%
  filter(pctFullVaccinated < 50)

# Get the Geography data for Maryland counties
counties <- counties() 
md_counties <- counties %>%
  filter(STATEFP == "24")

# Get the Zip Code data 
zctas <- zctas()

# Join the Zip Code geography data with the 
md_zctas <- md_by_zip_lt50 %>% left_join(zctas,  by=c("ZIP_CODE"="ZCTA5CE10"))

mapTitle <- "Maryland Zip Codes with fewer than 50% fully vaccinated"

# Create a sall dataframe with just the date field, although I could have used the variable I created earlier also
currentRec1 <- md_by_zip_lt50 %>% head(1) %>% select(vaccination_date)

# Subtitle shows the "as of" date
mapSub <- paste("As of ", currentRec1$vaccination_date, sep="")

# Plot the first layer of counties in white, then the Zips with <50% in orange
#
ggplot() + 
  geom_sf(data=md_counties, fill="white") + 
  aes(geometry = geometry) +
  geom_sf(data=md_zctas, fill="orange") + 
  aes(geometry = geometry) +
  theme_minimal() +
  labs(title = mapTitle,  
       subtitle = mapSub)

```
