---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
mykey <- scan("../../censusapikey.txt", what=character())
census_api_key(mykey, install=TRUE, overwrite=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
acs1 <- load_variables(2019, "acs5", cache = TRUE)
View(acs1)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file
removed_ppp_applications <- read_rds("data/lab_08.rds")

```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1**

```{r}
md <- get_acs(geography = "county",
              variables = c(medianage = "B01002_001"),
              state = "MD")
```

**Q2** Which Maryland counties have a median age of at least 50?
**A2**
Talbot and Worcester counties
```{r}
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3**
Prince Georges, followed by Montgomery, Caroline, Frederick, and Talbot
24033	Prince George's County, Maryland	36642	143125	0.256013974
24031	Montgomery County, Maryland	36099	177565	0.203300200
24011	Caroline County, Maryland	598	5832	0.102537723
24021	Frederick County, Maryland	3837	43949	0.087305741
24041	Talbot County, Maryland	449	5196	0.086412625
```{r}
# MD counties with 5-17 year olds 
# AGE BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 5 YEARS AND OVER Speak Spanish
# age 5-17
# Variable name is B16007_004
md1 <- get_acs(geography = "county",
              variables = c(youngSpanish = "B16007_004"),
              state = "MD")

md1 %>%
  select(GEOID, NAME, estimate) %>% rename(span5_17=estimate) -> md1

# MD counties with 5-17 year olds
# AGE BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 5 YEARS AND OVER
# Variable name is B16007_001
md2 <- get_acs(geography = "county",
              variables = c(young = "B16007_002"),
              state = "MD")
md2 %>%
  select(GEOID, NAME, estimate) %>% rename(all5_17=estimate) -> md2

youngdata <- md1 %>% inner_join(md2)
youngdata %>%
  mutate(pctage = span5_17/all5_17) %>%
  arrange(desc(pctage))
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4**
Well, 3 of the counties have a high overall population of 5-17 year olds, but the other 2 have relatively small populations of 5-17 year olds.
Garrett county have very few spanish-speaking at home 5-17 year olds. 

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to creeate separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5**
60620	had the most removed PPP applications with 165
96.4%
```{r}
# Clean the ZIP code
# group by and count the number of records

topZips <-
  removed_ppp_applications %>%
  mutate(zip5 = str_sub(zip, 1, 5)) %>%
  group_by(zip5) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

HighestZip <-topZips %>%
  slice(1) %>%
  select(zip5)

# BLACK OR AFRICAN AMERICAN ALONE OR IN COMBINATION WITH ONE OR MORE OTHER RACES
# B02009_001


blackByzcta <- get_acs(geography = "zcta",
              variables = c(black = "B02009_001"))
blackByzcta %>% 
  select(GEOID, NAME, estimate) %>% rename(blackpop=estimate) ->  blackByzcta

#
# All RACE is B02001_001
allByzcta <-get_acs(geography = "zcta",
              variables = c(pop= "B02001_001"))
allByzcta %>%
  select(GEOID, NAME, estimate) %>% rename(allpop=estimate) ->  allByzcta

popdata <- allByzcta %>% inner_join(blackByzcta)
popdata <- popdata %>%
  mutate(pctage = blackpop/allpop) %>%
  arrange(desc(pctage))

# select the row for the zip code with the highest # of removed PPP loans
popdata %>%
  filter(str_detect(GEOID, HighestZip$zip5))

```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6**
9 out of 10 are in the South side of Chicago. 1 is in Fultin COunty, GA (Atlanta, GA)

**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7**
WHo was the servicing_Lender, rural_urban_indicator, race, business type.