---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)


# Load janitor here
library(janitor)
```

## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)
ppp_wv_loans <- read_rds("lab_03.rds") %>%
  mutate(id = as.character(id))


```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia got the most loans (in number of loans)? How many did each get? To answer this question, you'll have to clean the very messy "city" field.

Note: To answer this question, please fully clean the city field for Kanawha County cities. Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1:
Charleston	2837			
South Charleston	484			
Saint Albans	304			
Dunbar	166			
Nitro	142	
```{r}
cleaned_wv_loans <- ppp_wv_loans %>%
  clean_names()
cleaned_wv_loans %>%
  filter(project_county_name == "KANAWHA") %>%
  mutate(city = str_to_title(city)) %>%
  group_by(city) %>%
  mutate(city = case_when(
    city == "Balle" ~ "Belle",
    city == "Charelston" ~ "Charleston",
    city == "Crosslanes" ~ "Cross Lanes",
    city == "Elkveiw" ~ "Elkview",
    city == "S Charleston" ~ "South Charleston",
    city == "So Charleston" ~ "South Charleston",
    city == "St.albans" ~ "Saint Albans",
    city == "St Albans" ~ "Saint Albans",
    city == "St. Albans" ~ "Saint Albans",
    TRUE ~ city
  )) %>%
  summarise(
    count=n()
  ) %>%
  #arrange(city)
  arrange(desc(count))
# Write your code in this codeblock.

```

Q2: What is the name of the business that got the largest loan (by dollar value) in the 5-digit ZIP code in West Virginia that had the highest dollar amount in total loans (by dollar value)? Write up a short paragraph that also answers these questions:

* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

A2: 
25301 is the 5-digit ZIP code in West Virginia that had the highest dollar amount in total loans. Zip 25301 is located in Charleston, and the business with the highest loan in that Zip code is JACKSON KELLY PLLC, a "full service law firm" in the amount $5,068,472.00. The forgiveness amount was $5,126,618.60, which is curiously greater than the loan amount. On the propublica web page for the company, there is a not that says "Includes any accrued interest." the difference between the forgiveness amount and the original loan amount is a little over 1%. Perhaps the forgiveness paid for the original amount plus accrued interest that had been allocate dor paid back to the bank over the time period?

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
top_wv_loans_by_zip <-
cleaned_wv_loans %>%
  mutate(city = str_to_title(city)) %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(amount = as.numeric(amount)) %>%
  group_by(zip) %>%
  summarise(
    total_amount=sum(amount)
  ) %>%
  arrange(desc(total_amount))
# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.
top_wv_loan <-
  cleaned_wv_loans  %>%
  mutate(city = str_to_title(city)) %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(amount = as.numeric(amount)) %>%
  filter (zip == "25301") %>%
  arrange(desc(amount))
```

Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  * What percentage of the total loan amount have been forgiven?
  * How did the U.S. House member vote on the legislation initially author zing the PPP loan program.

Feel free to add any interesting "color" (little interesting details) you find in your research.  

Some hints to help you answer this question:

* West Virginia's three U.S. House members are listed [here](https://www.govtrack.us/congress/members/WV#representatives)
* To find businesses they or their families own, you can search for financial disclosure forms for U.S. House members [here](https://disclosures-clerk.house.gov/PublicDisclosure/FinancialDisclosure#Search).
  * To speed things up, here are direct links to the three U.S. House members most recent financial disclosures, which I found by searching the site linked above.
  * [Rep. 1](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042042.pdf)
  * [Rep. 2](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10040330.pdf)
  * [Rep. 3](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042734.pdf)
* When you examine the disclosure forms, there may be lots of stock sales and retirement accounts mixed in.  Ignore those. In the table in the disclosure forms, look for "Partnership" or "C-Corp" in the column named "Income Type"
* This is a link to details on passage of the [CARES ACT](https://www.congress.gov/bill/116th-congress/house-bill/748). You can find a link to how the House voted by clicking on a link from [this page](https://www.congress.gov/bill/116th-congress/house-bill/748/actions).

A3:
There are 5 businesses that are associated with representative Carol Devine Miller:
DM MOTOR, INC
DUTCH MILLER OF CHARLESTON, INC
DUTCH MILLER OF CHARLOTTE, INC
DUTCH MILLER SUBARU, INC
DUTCH MILLER CHEVROLET, INC.

These are all automobile dealerships. The total of the loan amounts was $3,105,300 and the totla amount forgiven was $3,140,471. The amount forgiven was 101.132% of the total of the loan amounts. Representative Miller was a cosponsor of the bill and voted for the bill.
```{r}

# Hint
# First, create a dataframe that only has loans to those five companies
# GGM Consulting LLC
# Maxwell Centre, LLC
# DEaB LLC
# DEaB II LLC
# 1088 Parker Dr.
# DM Motors
# Dutch Miller of Charleston
# Dutch Miller Subaru
# MFM Realty
#
rep_wv_loans <-
cleaned_wv_loans  %>%
  mutate(city = str_to_title(city)) %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(amount = as.numeric(amount)) %>%
  mutate(name = str_to_upper(name)) %>%
  filter(str_detect(name, "DUTCH |GGM|DM MOT"))


# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies
rep_wv_loans %>%
  summarise(total_amount=sum(amount), total_forgiven = sum(forgiveness_amount))

```
