---
title: "lab_05"
author: "Sean Mussenden"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(janitor)

```

## Load Data

You'll need to load three data sets for this:

* The West Virginia slice of the PPP loan data (lab_05.rds).
* A "lookup table" that allows you to translate NAICS (industry) numeric codes to industry titles (naics_codes.csv).
* A table of West Virginia population by county (American Community Survey, 2019 5-year averages) (wv_population_county.csv).

All three data sets are in the data folder.  Write code to load the three in the codeblock below.

**Task** Create a codeblock below this task, then read the data sets in in and assign them to appropriate variable names. There's a fourth data set you'll use in this lab, too, of selected loans in ZIP Code 25401. But there's no need to load it now.

## Answer questions

**Q1.** In the data folder, there is a csv called zip_25401_loan_sample.csv.  It contains a sample of loans from West Virginia ZIP Code 25401. 

As we read earlier this semester, [multiple loan applications coming from multiple businesses at the same residential street address](https://www.nytimes.com/2021/08/17/business/ppp-fraud-covid.html) might point to fraud. Or it could alert us to companies that used [multiple corporate entities to get more money than envisioned](https://www.propublica.org/article/different-names-same-address-how-big-businesses-got-government-loans-meant-for-small-businesses) by the original law that authorized the program.   

You are going to examine this data to see if we can find a lot of loans coming from the same address.  Here's the problem: the street address field is pretty messy.  The same address appears with minor variations --  "1003 Sushruta Dr" vs "1003 SUSHRUTA DR" -- that will prevent proper grouping. 

First, upload the data into Open Refine and standardize/clean the address field. If you've done it properly, you should have 65 discrete addresses. 

Then export the data from Open Refine, and move it to the proper folder. 

Next, load it in the codeblock below, assigning it to an appropriate variable name. 

Then answer these questions:
* What is the street address in this data that has the most loans?
* How many loans are there at that street address?
* What are the names of the businesses at that address?

**A1.**
126 E Burke St, Martinsburg WV, 6 loans at that address:
DREW HOLDINGS LLC				
BRIX27, LLC				
ABRAHAM ASHTON				
HUB CO-OP LLC				
HUB CO-OP, LLC				
RONIN PROPERTIES LLC
```{r}
# Load the WV loan data
ppp_wv_loans <- read_rds("data/lab_05.rds") %>%
  mutate(id = as.character(id))
# Load the NAICS data
naics_codes <- read_csv('data/naics_codes.csv')
# Load the WV ACS data
wv_acs_population <- read_csv("data/wv_population_county.csv")
# Load the 25401 cleansed data
wv_loans_25401 <- read_csv("data/zip_25401_loan_sample-csv.csv")

# * What is the street address in this data that has the most loans?
# * How many loans are there at that street address?
# * What are the names of the businesses at that address?
wv_loans_25401 %>%
  group_by(clean_address) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))

wv_loans_25401 %>%
  filter(str_detect(clean_address, "126 E Burke St")) %>%
  select(name)

```

**Q2.** Do some web research on the businesses that answered question 1.  

Google the street address.  Google the business names and search for their corporate records on [Open Corporates](https://opencorporates.com/). Be sure to find the website of the name of the company that appears twice in the list of businesses and develop an understanding of what it does. 

Based on your research, does it seem suspicious that this collection of businesses all got loans using the same address? Why or why not. Use specific facts identified in your research to support your position. 

**A2.**
The Hub is a cooperative office space, basically a rental space for businesses. It also runs a cafe at the same location. Brix27 is a wine bar that is *not* located at 126 E Burke St. There are two real estate LLCs. Robert Johnson and Abraham Ashton are associated with all of these organizations. Abraham Ashton also has a law office at 208 S Queen St, around the corner from the 126 E Burke location. Not really suspicious - looks like a partnership between Ashton and Johnson to create a business office, and they register their businesses at that location.

**Q3.** Start by using the West Virginia slice of the PPP loan data that you loaded at the start of the lab to create a subset of PPP loans in West Virginia's second largest county (which you can find in wv_population_county). And then use that table you created to answer the following questions:

* Which city in that county had the highest number of loans? 
* In that city, which industry title had more loans than any other industry title? 

Requirement: you MUST export a dataframe of PPP loans from R Studio at some point in the process (not necessarily at the beginning!), load it into Open Refine, clean the city column, export it from Open Refine, and reimport into R Studio. To export data, you will use the write_csv() function.

Guidance: there are a lot of steps you'll need to take to answer this question. You may or may not find it helpful to write out in English what you plan to do step-by-step before you start writing code.   

**A3.**
Martinsburg had the highest number of loans.
Full-Service Restaurants is the most frequently appearing title. According to Wikipedia, Martinsburg is the largest city in the Eastern Panhandle of WV and is the county seat of Berkeley county. 
```{r}
# Sort counties by population
wv_acs_population %>%
  arrange(desc(population_2019))
# Find cities in that county
wv_berkeley_to_export <- ppp_wv_loans %>%
  filter(str_detect(project_county_name, "BERKELEY"))
  
# Export to cleanse
write_csv(wv_berkeley_to_export, "data/wv_berkeley.csv")
# Reimport
wv_berkeley_clean <- read_csv("data/wv_berkeley-csv.csv")
#, group and and sort by loan count
wv_berkeley_clean %>%
  group_by(clean_city) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
# JOIN with the NAICS codes
wv_berkeley_clean_w_naics <- wv_berkeley_clean %>% left_join(naics_codes)
# For that city, group and sort by industry title
wv_berkeley_clean_w_naics %>%
  filter(str_detect(clean_city, "Martinsburg")) %>%
  group_by(title) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
```

**Q4.** What are your two best hypotheses, which you'd need to confirm with reporting and further analysis, that explain why that industry is at the top of the list?

**A4.** 
As the "most-happening" city in the area, as well as effectively being the center for local businesses (dentists, physicians, beauty salons, brokerages, religious organizations, etc.), it is most likely to have highest "transient" population as well as largest population. ALl those people have to eat, so it is likely to be a magnet for restaurants.

**Q5.** Start with a table of loans to all businesses in the city and industry that answered question 3. Answer the following questions:
* What is the name of the business that got the highest approved loan amount? 
* How much was it for?
* When was it approved?
* How many jobs does the data say were retained?
* Is there a difference between the business' name in the PPP data and the name its customers know it by? If so, what is that name?
* How many locations does this business have? 
* Did one of its locations close during the pandemic, either before or after it got the loan?

Hint: you will not find the answers to the last three questions in the data.  You could call them directly to get that information, but I don't want you to do that for this assignment.  Instead, do some web research. I would start by Googling the company name from the data and looking at the page that comes up for the business from at http://apps.sos.wv.gov/. I would use information I found on that page and use info about the company from Google, the [Wayback machine](https://archive.org/web/) (which lets you look at older versions of a company's website), Yelp, and Facebook. 

**A5.**
COTTLE CHRISTI L LLC got the highest amount.
The loan was for $280,434.00.
The loan was approved 2021-02-17.
It says that there were 94 jobs retained.
Others know this business as "Kitzie's Restauarant and Lounge."
According to the website, (http://www.kitziesrestaurant.com/) there are 2 locations, Martinsburg and Spring Mills. 
After looking at Yelp reviews for both locations I go tthe idea that both were fully open (and packed) during the pandemic. I looked at the wayback machine snapshots of their web sites, and one snapshot in October 2020 from the Spring Mills location said "Kitzie's is Excited to Announce that We are Open for Indoor Dining!" No changes on the other location's web site. So it may be a case where a business applied for  aloan but really did not need it?
``` {r}
wv_martinsburg_restaurants <-
  wv_berkeley_clean_w_naics %>%
  filter(str_detect(clean_city, "Martinsburg")) %>%
  filter(str_detect(title, "Full-Service Restaurants"))
  
```