---
title: "lab_04"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
# add code here to accomplish the task
maryland_ppp <- read_rds("lab_04.rds")

```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the book chapter. There's a particular set of steps you'll need to perform to make that work properly.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
# add code here to accomplish the task
naics_codes <- read_csv('data/naics_codes.csv')

```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.

```{r}
# add code here to accomplish the task
maryland_ppp_with_naics <- maryland_ppp %>% left_join(naics_codes)
maryland_ppp_with_naics <- maryland_ppp_with_naics %>% mutate(zip5 = str_sub(zip, 1, 5))

```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one:

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character.

```{r}
# add code here to accomplish the task
maryland_zcta <- read_csv('data/maryland_zcta.csv')
maryland_zcta <- maryland_zcta %>% mutate(across(ZCTA5N, as.character))

```

Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
# add code here to accomplish the task
maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics %>% left_join(maryland_zcta, by=c("zip5"="ZCTA5N"))
```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?
**A1** [write your answer here]
Taxi Service	7363		
Beauty Salons	7275		
Full-Service Restaurants	5490		
Offices of Real Estate Agents and Brokers	4443		
All Other Personal Services	3956	
```{r}
# Write the code to produce your answer here
maryland_ppp_with_naics_and_demographics %>%
  group_by(title) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** [write your answer here]
Full-Service Restaurants	1488		
Offices of Physicians (except Mental Health Specialists)	580		
Plumbing, Heating, and Air-Conditioning Contractors	458		
Offices of Dentists	419		
Commercial and Institutional Building Construction	402	
```{r}
# Write the code to produce your answer here
maryland_ppp_with_naics_and_demographics %>%
  filter(amount>150000) %>%
  group_by(title) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** [write your answer here]
zip5  total     count
20850	474525058	3186		
20878	151646470	2188		
21043	72404137	1385		
20876	43039899	723		
20871	37762652	734		
```{r}
# Write the code to produce your answer here
# See https://geodata.md.gov/imap/rest/services/Demographics/MD_CensusData/FeatureServer/1
# PNHA: Percent Non Hispanic Asian Persons
# Group by zip5
# get the total and count by summarise
# Order in descending order by sum

maryland_ppp_with_naics_and_demographics %>%
  filter(PNHA >= 20) %>%
  group_by(zip5) %>%
  summarise(
    total = sum(amount),
    count = n()
    ) %>%
  arrange(desc(total))

```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4** [write your answer here]
20850	28994798.00	240	

ZCTA5N PNHA
20850	 23.5	
```{r}
# Write the code to produce your answer here
# race == "Asian"
# Group by zip5
# get the total and count by summarise
# Order in descending order by sum
maryland_ppp_with_naics_and_demographics %>%
 filter(race=="Asian") %>%
  group_by(zip5) %>%
  summarise(
    total = sum(amount),
    count = n()
    ) %>%
  arrange(desc(total))

# Select Non-Hispanic Asian percentage from maryland_zcta
maryland_zcta %>%
  filter(ZCTA5N=="20850") %>%
  select(ZCTA5N,PNHA)

```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants?
**A5** [write your answer here]
Missing data. >80% of the applications have "Unanswered" as the value for the race field. So most applicants did not report their race.

**Q6** Find any Maryland zip codes that have loans where the PPP data lists the applicants as being Asian but where the zip code does not appear to have Asian population in the ZCTA data? What do you think is going on here?
**A6** [write your answer here]
zip5  count
21031	6			
20701	4			
21240	3			

For 21031, the total population is 4 people, probably the businesses are run by people who reported their race as Asian but they do not *live* in that Zip code
For 20701, total population is 2 people, so probably the same.
For 21240, that is the Zip code for BWI airport. The total population is 0. There may be businesses being run within the compound of the airport, but no one lives there.
```{r}
# Write the code to produce your answer here
# Get counts by race
maryland_ppp_with_naics_and_demographics %>%
 filter(zip5=="20850") %>%
  group_by(race) %>%
  summarise(
    count = n()
    ) %>%
  arrange(desc(count))
#
# look for where Asian population percentage is 0?
#
maryland_ppp_with_naics_and_demographics %>%
  filter(race =="Asian") %>%
  filter(PNHA == 0) %>%
  select(zip5) %>%
  group_by(zip5) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?
**A7** [write your answer here]
There are 44 industries with just 1 loan. Certainly, I would be curious about "Ammunition (except Small Arms) Manufacturing" with a loan exceeding $1M. ALso, the same company got a second PPP loan but used a different NAICS code for "Small Arms, Ordnance, and Ordnance Accessories Manufacturing".
```{r}
# Write the code to produce your answer here
maryland_ppp_with_naics_and_demographics %>%
  group_by(title) %>%
  summarise(
    total = sum(amount),
    count = n()
    ) %>%
  arrange(count) %>%
  filter(count == 1) %>%
  arrange(desc(total))
```
